Gender Dynamics in Academia: Analyzing Regrade Requests and Course Load Trends, 2010-2016

### 2024-02-20 outputs/paper/paper.qmd
Updates

---
title: "Gender Dynamics in Academia: Analyzing Regrade Requests and Course Load Trends, 2010-2016"
subtitle: "A replication of Li and Zahar;"
author: Faiza Imam, Quang Mai & Catherine Punnoose 
thanks: "Code and data are available at: https://github.com/ponolite/gender_grade.git, DOI: 10.1257/pol.20210053"
date: today
date-format: long
abstract: "Institutionalized and traditional, academia often exacerbates gender differences. This paper then replicates the gender implications behind regrade requests of undergraduate students in Colorado State University from 2010-2016. By analyzing the number of classes taken by students, their willingness to request regrade and actual grade changes, this paper explores fluctuations in regrade request patterns throughout the semester. Our results show that on average males students are more likely to submit regrade requests, and-even when they request regrade for more courses-still have better grade changes compared to female students. We thus call for further investigation around this topic to understand gender disparities within the world of academia."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Load data
inst_data <- read.csv(here::here("inputs/data/inst_data_cleaned.csv"))
std_data <- read_csv(here::here("inputs/data/stdsurvey.csv"))
```

# Introduction

It’s a heteronormative norm that we consider men to be assertive and leaderly, while women are nurturing and supportive. Often, women who choose to break out of these stereotypes and speak out against misogyny are labelled ‘emotional’ or ‘aggressive’, hindering their professional mobility [@AACSB].  Even in academic settings, women tend to be penalized when achieving better grades. A study conducted by Natasha Quadlin, suggests that due to traditional, gendered stereotypes, women’s academic achievements are often disregarded compared to their male counterparts when both enter the labour market. Men of ‘high-achieving’ academic backgrounds tend to get favourable callbacks for interviews compared to their female counterparts within the same field [@SageJournals].

This begs the question of whether these unequal tendencies exist in higher education for women, attempting to have their academic achievements recognized and improved by regrade requests. Many reasons can account for this gendered inequality. First, being subjected to societal norms, women tend to be more non-confrontational, resulting in them accepting their academic results as they are [@AACSB]. Second, with differing confidence levels and personalities, women tend to second-guess their abilities, hence preventing them from requesting aid from their institutions as they see the entire process as ‘tedious’. Lastly, fear of rejection, women tend to harbour higher levels of stress and embarrassment when considering regrade requests as they assume that the result will often be a rejection [@SageJournals]. A paper published by the American Journal of Political Science by Li and Zahar delves into the differences that arise between genders when requesting grade changes among college students from Colorado State University from 2010 to 2018. According to their findings, male students are more likely to request grade changes than their female counterparts, accounting for a whopping 18.6% higher of receiving a favourable grade change. Regardless of the class number, males are more likely to ask for grade changes throughout the semester [@InsideHigherEd]. Overall there is a gender disparity among college students regarding regrade requests. Hence, by examining biases inherent within higher education, the paper contributes to understanding the power dynamics between two gender demographics.

Our paper will reproduce, ‘Ask and You Shall Receive? Gender differences in Regrades in College’ by Li and Zafar. Accordingly, our estimand asserts that upon a regrade request, males are more likely to receive a favourable grade change regardless of the time during the semester, and regardless of the number of classes that they have. Our study attempts to pose and answer these questions, ‘Does gender bias influence regrade requests in college settings? Are there fluctuations in regrade request patterns among male and female students regarding their class number?’ Structurally, our paper starts with a brief dissection of the original paper’s findings and our own interpretation of the data. Subsequently, we analyze the results garnered from the data with suitable tables, and double stacked bar charts. Finally, to conclude, we will discuss our findings, entailed by an analysis of the data’s loopholes, ethical biases and future research implications.

@@

Published by the ‘American Economic Journal: Economic Policy’, this paper reproduced the paper ‘Ask and You Shall Receive? Gender Differences in Regrades in College’ and its replication package published on `openICPSR` [cite].  This data package conveys information related to survey data from instructors and students on the pervasiveness of regrade requests based on gender identities from 2010 to 2018. Particularly, ‘instsurvey.dta’ contains all 190 observations of actual regrade requests in 2018, along with 61 variables concerning the nature of each regrade request and whether or not each instance initiated grade changes. On the other hand, ‘stdsurvey.dta’ conveys all 3885 observations of student surveyees who have contemplated requesting regrades, with 52 variables detailing instances where they didn’t initiate such requests. With these two datasets representing two perspectives on regrade requests, there is much power dynamic to unpack, helping this paper assess the gendered nature of regrade requests in the undergraduate setting.

@@

Most prominent to our reproduction lies in the tactile grade changes of regrade requests based on each student’s gender identity. As such, we first simplified ‘instsurvey.dta’, which featured 190 observations. We cleaned the data by selecting relevant variables such as “maleup_fn”, “malesame_fn”, “maledown_fn”, “femaleup_fn”, “femalesame_fn”, “femaledown_fn”, “maleup_mt”, “malesame_mt”, “maledown_mt”, “femaleup_mt”, “femalesame_mt”, and “femaledown_mt”. The respective variables provide data on the number of male and female students whose grades increased, stayed the same or decreased both during and at the end of the semester. We further modified the data by calculating the mean of each variable using the mean() function. The mean is needed to determine the percentage of male or female students who received a grade increase, decrease, or no change. A new data frame is then created per focus area. In total four graphs were created to show the percentage of male and female students who received a grade increase, decrease, or no change after requesting for a regrade both during and at the end of the semester.

@@

@@

classes_dta <- std_data |>
  select(std_male, num_class, consider_regrade)|>
  rename(Gender = std_male) |>
  filter(is.na(consider_regrade) | consider_regrade != "No")

classes_dta <- classes_dta |>
  group_by(num_class, Gender, consider_regrade) |>
  count() |>
  group_by(Gender, consider_regrade) |>
  mutate(Percentage = round(n/sum(n)*100, 1))|>
  mutate(num_class = case_match(num_class, 
                                0 ~ "0 class",
                                1 ~ "1 class",
                                2 ~ "2 classes",
                                3 ~ "3 classes",
                                4 ~ "4 classes",
                                5 ~ "5 classes"))

@@

classes_dta |>
  filter(consider_regrade != "No") |>
  rename("Count" = n, "Number of Classes" = num_class, "Consider Regrade" = consider_regrade) |>
  head(12) |>
  kable(
    col.names = c("Number of Classes", "Gender", "Consider Regrade", "Count", "Percentage"),
    booktabs = TRUE,
    align = c("c", "c", "c", "c", "c")
  )

@@

Conversely, the ‘stdsurvey.dta’ dataset featured 3885 observations of all potential regrade requests, which provided us with insights into gender identity and student psychology when approaching regrades. To simplify the data, we first clean its names and select relevant variables, including ‘std_male’, ‘consider_regrade’, ‘num_class’ and ‘participantcode’ which respectively convey data on students’ gender, whether or not they have considered regrading, the number of classes they have considered regrades and identification. For ‘std_male’, a ‘1’ signifies a male student while a ‘0’ signifies a female student; in which, to clean the dataset, we’ve renamed the column to ‘gender’ and modified ‘1’ to be ‘Male’ while ‘0’ to be ‘Female’. To garner data on students who considered regrading, we used `filter` to get rid of observations with ‘No’ for their ‘consider_regrade’ column. `mutate` was also leveraged to generate a column on ‘percentage’ of students who considered regrading ([@tbl-genderclass]). Proceedingly, we used the cleaned data to generate a double-stacked bar chart of the percentage of female and male students who considered regrading for each of the categories of the number of classes.

# Results

@@

```{r}
#| label: fig-RegreadeResultsByTime
#| fig-cap: Instructor Survey Regrade Results by Timing and Students’ Gender
#| fig-subcap: [Panel A. End of semester male students, Panel B. End of semester female students, Panel C. During semester male students, Panel D. During semester female students]
#| layout-ncol: 2
#| warning: false
#| message: false
#| echo: false

#Code Assisted from Chat GPT 3.5
# End of Semester Results MALE

male_dta <- inst_data |>
              select(maleup_fn, malesame_fn, maledown_fn)

maleup_avg <- mean(male_dta$maleup_fn, na.rm = TRUE)
malesame_avg <- mean(male_dta$malesame_fn, na.rm = TRUE)
maledown_avg <- mean(male_dta$maledown_fn, na.rm = TRUE)

males <- data.frame(
  value = c(maleup_avg, malesame_avg, maledown_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

@@

# Create a bar graph using ggplot
ggplot(males, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

# End of Semester Results FEMALE

female_dta <- inst_data |>
              select(femaleup_fn, femalesame_fn, femaledown_fn)

femaleup_avg <- mean(female_dta$femaleup_fn, na.rm = TRUE)
femalesame_avg <- mean(female_dta$femalesame_fn, na.rm = TRUE)
femaledown_avg <- mean(female_dta$femaledown_fn, na.rm = TRUE)

females <- data.frame(
  value = c(femaleup_avg, femalesame_avg, femaledown_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

@@

# Create a bar graph using ggplot
ggplot(females, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

# During Semester Results MALE

male_mt_dta <- inst_data |>
              select(maleup_mt, malesame_mt, maledown_mt)

maleup_mt_avg <- mean(male_mt_dta$maleup_mt, na.rm = TRUE)
malesame_mt_avg <- mean(male_mt_dta$malesame_mt, na.rm = TRUE)
maledown_mt_avg <- mean(male_mt_dta$maledown_mt, na.rm = TRUE)

males_mt <- data.frame(
  value = c(maleup_mt_avg, malesame_mt_avg, maledown_mt_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

@@

# Create a bar graph using ggplot
ggplot(males_mt, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

# During Semester Results FEMALE

female_mt_dta <- inst_data |>
              select(femaleup_mt, femalesame_mt, femaledown_mt)

femaleup_mt_avg <- mean(female_mt_dta$femaleup_mt, na.rm = TRUE)
femalesame_mt_avg <- mean(female_mt_dta$femalesame_mt, na.rm = TRUE)
femaledown_mt_avg <- mean(female_mt_dta$femaledown_mt, na.rm = TRUE)

females_mt <- data.frame(
  value = c(femaleup_mt_avg, femalesame_mt_avg, femaledown_mt_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

@@

# Create a bar graph using ggplot
ggplot(females_mt, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

@@

Suceedingly, @fig-RegradeByClass shows the averaged distribution of the number of classes in which students are considered for regrades by gender. Observable in the graph, there is only data available on students who considered regrades in ‘0 class’ to ‘5 classes or more’ ([@fig-RegradeByClass]). Therefore, with the number of classes limited to 5, it’s more than likely that surveys understate the regraded data [@GenderDiff]. Observably, with no or 0 classes, the majority of students didn’t consider regrade requests, with females upholding 41.2 % while males upholding 39.3%, making it a 1.9% difference. This trend of females considering for regrades is higher than males when the class number is lower, where at 1 class, a whopping 30.4% of females are considered for regrades in contrast to a meagre 22.7% of males. This translates to a huge margin of 7.7% between the two genders. However, as the number of classes extended above 1, or from 2 to 5 or more classes, more males considered asking for a regrade. For instance, at 3 classes, 8.9% of males considered for regrades as opposed to 7.1% of females; at 4 classes, 3.3% of males considered for regrades compared to 2% of females and finally, at 5 classes or more, a great 5.2% of males considered for regrades compared to a small 2.1% females. Statistically then, while the difference is negligible in some categories of class number, males are more likely to consider and ask for regrades compared to females (except for those with 1 class).

@@

# Discussion

The original paper looked to analyze the gender difference in a college setting upon regrade requests. The main questions Li and Zahar analyzed were whether male students are more frequent in their regrade requests compared to their female counterparts. The authors also look to analyze whether these requests result in favorable outcomes for either gender.  By replicating this paper we look to understand the gender differences in favourable regrade requests with factors of course load and request frequencies throughout the semester.

@@

One key point to consider is the disparity in outcomes between male and female students when requesting a regrade. The data shows that a higher percentage of male students (62%) received no change in their grade after requesting a regrade during the semester compared to female students (59%). This explains the notion that males typically send out higher requests of regrades compared to their female counterparts. This could be attributed to the social implications placed on each gender, confirming the concept that men are more assertive hence their inclination to ask for regrades. Women are more accepting hence they refrain from attempting to submit a request. When looking at the percentage of students who did receive an increase in their grade after requesting a regrade, it turns out that a higher proportion of female students (40%) experienced this outcome compared to male students (37%). This suggests that, although male students were less successful overall in achieving grade changes when changes did occur, they were slightly more likely to increase for female students. Statistically, this 3% difference in outcomes, does not explain a great deal regarding the skewed gender implications but further research is required to understand the broader implications.

>>>>

---
title: "Gender Dynamics in Academia: Analyzing Regrade Requests and Course Load Trends, 2010-2016"
author: Faiza Imam, Quang Mai & Catherine Punnoose 
thanks: "Code and data are available at: https://github.com/ponolite/gender_grade.git, DOI: https://doi.org/10.48152/ssrp-bjze-cn68"
date: today
date-format: long
abstract: "Academia often exacerbates gender differences. This paper further explores the gender implications behind regrade requests of undergraduate students at Colorado State University from 2010-2016. By analyzing changes in regrade request patterns throughout the semester, actual grade changes, and instructor’s gender identity, our results show that on average male students are more likely to submit regrade requests, and are more likely–even when they request regrade for more courses–to have better grade changes compared to female students. In addition, extrapolation on the original study seems to suggest that both female and male instructors are more likely to give out favourable regrades for female students, which raises interesting questions on gender dynamics and regrading in the college setting."
format: pdf
fig-pos: 'H'
toc: true
number-sections: true
bibliography: references.bib
---

@@

# Load data
inst_data <- read.csv(here::here("inputs/data/inst_data_cleaned.csv"))
std_data <- read_csv(here::here("inputs/data/stdsurvey.csv"))
inst_data_cleaned <- read_csv(here::here("inputs/data/inst_data_cleaned.csv"))
```

# Introduction {#sec-intro}

It’s a societal norm that we consider men to be assertive and dominant in leadership roles while women are nurturing and supportive. Often, women who choose to break out of these stereotypes and speak out against misogyny and sexism are labelled ‘emotional’ or ‘aggressive’, hindering their professional mobility [@AACSB].  Even in academic settings, women tend to be penalized when achieving better grades and opportunities. A study conducted by Natasha Quadlin, suggests that due to traditional, gendered stereotypes, women’s academic achievements are often disregarded compared to their male counterparts when both enter the labour market. Men of ‘high-achieving’ academic backgrounds tend to get favourable callbacks for interviews compared to their female counterparts with the same credentials [@SageJournals].

This begs whether these unequal tendencies exist in higher education for women, as they attempt to have their academic achievements recognized by requesting regrades. There are many reasons to account for when analyzing gendered inequality. First, being subjected to societal norms, women tend to be more non-confrontational, resulting in them accepting their academic results as they are instead of fighting for them [@AACSB]. Second, with differing confidence levels and personalities, women tend to second-guess their abilities, hence preventing them from requesting aid from their institutions as they see the entire process as ‘tedious’ and “unnecessary’ [@SageJournals]. Lastly, fear of rejection, women tend to harbour higher levels of stress and embarrassment when considering regrade requests as they assume that the result will often be a rejection from their instructors [@SageJournals]. A paper published by the American Journal of Political Science by Li and Zahar delves into the differences that arise between genders when requesting grade changes among college students from Colorado State University from 2010 to 2018. According to their findings, male students are more likely to request grade changes than their female counterparts, accounting for 18.6% higher of receiving a favourable grade change. Regardless of the class number, males are more likely to ask for grade changes throughout the semester [@InsideHigherEd]. Overall there is a gender disparity among college students regarding regrade requests. Hence, by examining biases inherent within higher education, the paper contributes to understanding the power dynamics between two gender demographics. In addition, to extrapolate from the original paper, our paper combines the student regrade data with instructors’ gender to see if there is gender dynamics between the two.

Thus, from the paper, ‘Ask and You Shall Receive? Gender differences in Regrades in College’ by Li and Zafar, we hope to look into the estimand that upon a regrade request, whether or not an instructor’s gender identity–when interacted with a student’s gender identity– affects the outcome of such a student’s re-grading. Our study attempts to pose and answer these questions, ‘Does gender bias influence regrade requests in college settings?’, ‘Are there fluctuations in regrade request patterns among male and female students regarding their class number?’ and ‘Is there a correlation between instructor's gender and the regrading process?’

Structurally, in @sec-intro, our paper starts with a brief dissection of the original paper’s findings and our interpretation of the data. Subsequently, in @sec-data and @sec-results, we analyze the results garnered from the data with suitable tables and double-stacked bar charts. Finally, to conclude, we will discuss our findings in @sec-discussion, which entailed an analysis of the data’s loopholes, ethical biases and future research implications.

@@

Published by the ‘American Economic Journal: Economic Policy’, this paper reproduced the paper ‘Ask and You Shall Receive? Gender Differences in Regrades in College’ and its replication package published on `openICPSR` [@openICPSR].  This data package conveys information related to survey data from instructors and students on the pervasiveness of regrade requests based on gender identities from 2010 to 2018. Particularly, ‘instsurvey.dta’ contains all 190 observations of actual regrade requests in 2018, along with 61 variables concerning the nature of each regrade request and whether or not each instance initiated grade changes. On the other hand, ‘stdsurvey.dta’ conveys all 3885 observations of student surveyees who have contemplated requesting regrades, with 52 variables detailing instances where they didn’t initiate such requests. With these two datasets representing two perspectives on regrade requests, there is much power dynamic to unpack, helping this paper assess the gendered nature of regrade requests in the undergraduate setting.

@@

Most prominent to our reproduction lies in the tactile grade changes of regrade requests based on each student’s gender identity. As such, we first simplified ‘instsurvey.dta’, which featured 190 observations. We cleaned the data by selecting relevant variables such as “maleup_fn”, “malesame_fn”, “maledown_fn”, “femaleup_fn”, “femalesame_fn”, “femaledown_fn”, “maleup_mt”, “malesame_mt”, “maledown_mt”, “femaleup_mt”, “femalesame_mt”, “femaledown_mt” and “inst_gender”. The respective variables provide data on the number of male and female students whose grades increased, stayed the same or decreased both during and at the end of the semester along with the instructor’s gender in each case. We further modified the data by calculating the mean of each variable using the mean() function. The mean is needed to determine the percentage of male or female students who received a grade increase, decrease, or no change. A new data frame is then created per focus area. In total four graphs were created to show the percentage of male and female students who received a grade increase, decrease, or no change after requesting for a regrade both during and at the end of the semester.

Further, to analyze the differences in favourable regrades between male and female instructors, we renamed the original variable, “inst_female”  into “inst_gender”. We also changed “0” into “Male” and “1” into “Female to clarify the dataset. After that, we attempt to create four bar graphs to assess the correlation between instructors’ gender identity and favourable grade increases.

@@

Conversely, the ‘stdsurvey.dta’ dataset featured 3885 observations of all potential regrade requests, which provided us with insights into gender identity and student psychology when approaching regrades. To simplify the data, we first clean its names and select relevant variables, including ‘std_male’, ‘consider_regrade’, ‘num_class’ and ‘participantcode’ which respectively convey data on students’ gender, whether or not they have considered regrading, the number of classes they have considered regrades and identification. For ‘std_male’, a ‘1’ signifies a male student while a ‘0’ signifies a female student; in which, to clean the dataset, we’ve renamed the column to ‘gender’ and modified ‘1’ to be ‘Male’ while ‘0’ to be ‘Female’. To garner data on students who considered regrading, we used `filter` to get rid of observations with ‘No’ for their ‘consider_regrade’ column. `mutate` was also leveraged to generate a column on ‘percentage’ of students who considered regrading ([@tbl-genderclass]). Proceedingly, we used the cleaned data to generate a double-stacked bar chart of the percentage of female and male students who considered regrading for each category of the number of classes.

@@

# Selecting required data 
classes_dta <- std_data |>
  select(std_male, num_class, consider_regrade)|>
  rename(Gender = std_male) |>
  filter(is.na(consider_regrade) | consider_regrade != "No")

# Cleaning classes_dta 
classes_dta <- classes_dta |>
  group_by(num_class, Gender, consider_regrade) |>
  count() |>
  group_by(Gender, consider_regrade) |>
  mutate(Percentage = round(n/sum(n)*100, 1))|>
  mutate(num_class = case_match(num_class,     # mutating data so it is easier to understand 
                                0 ~ "0 class",
                                1 ~ "1 class",
                                2 ~ "2 classes",
                                3 ~ "3 classes",
                                4 ~ "4 classes",
                                5 ~ "5 classes"))

@@

# Creating a Table of classes_dta
classes_dta |>
  filter(consider_regrade != "No") |>
  rename("Count" = n, "Number of Classes" = num_class, "Consider Regrade" = consider_regrade) |>
  head(12) |>
  kable(
    col.names = c("Number of Classes", "Gender", "Consider Regrade", "Count", "Percentage"),
    booktabs = TRUE,
    align = c("c", "c", "c", "c", "c")
  )

@@

# Results {#sec-results}

@@

```{r}
#| label: fig-RegreadeResultsByTime
#| fig-cap: Instructor Survey Regrade Results by Timing and Students’ Gender
#| fig-subcap: [Panel A. End of semester male students, Panel B. End of semester female students, Panel C. During semester male students, Panel D. During semester female students]
#| layout-ncol: 2
#| warning: false
#| message: false
#| echo: false
#| fig-pos: "t"

#Code Assisted from Chat GPT 3.5

# FIGURE 1A: End of Semester Results MALE Bar Graph

# Selecting data on final grades for male students 
male_dta <- inst_data |>
              select(maleup_fn, malesame_fn, maledown_fn)

# Calculating average of grade increase, no change, and decrease for male students 
maleup_avg <- mean(male_dta$maleup_fn, na.rm = TRUE)
malesame_avg <- mean(male_dta$malesame_fn, na.rm = TRUE)
maledown_avg <- mean(male_dta$maledown_fn, na.rm = TRUE)

# Creating new frame with selected data 
males <- data.frame(
  value = c(maleup_avg, malesame_avg, maledown_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

@@

# Create a bar graph using ggplot
ggplot(males, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent", fill = "Category") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

# FIGURE 1B: End of Semester Results FEMALE Bar Graph

# Selecting data on final grades for female students 
female_dta <- inst_data |>
              select(femaleup_fn, femalesame_fn, femaledown_fn)

# Calculating average of grade increase, no change, and decrease for female students 
femaleup_avg <- mean(female_dta$femaleup_fn, na.rm = TRUE)
femalesame_avg <- mean(female_dta$femalesame_fn, na.rm = TRUE)
femaledown_avg <- mean(female_dta$femaledown_fn, na.rm = TRUE)

# Creating new frame with selected data 
females <- data.frame(
  value = c(femaleup_avg, femalesame_avg, femaledown_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

@@

# Create a bar graph using ggplot
ggplot(females, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent", fill = "Category") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

# FIGURE 1C: During Semester Results MALE Bar Graph

# Selecting data of grades during the semester for male students 
male_mt_dta <- inst_data |>
              select(maleup_mt, malesame_mt, maledown_mt)

# Calculating average of grade increase, no change, and decrease for male students 
maleup_mt_avg <- mean(male_mt_dta$maleup_mt, na.rm = TRUE)
malesame_mt_avg <- mean(male_mt_dta$malesame_mt, na.rm = TRUE)
maledown_mt_avg <- mean(male_mt_dta$maledown_mt, na.rm = TRUE)

# Creating new data frame with selected data
males_mt <- data.frame(
  value = c(maleup_mt_avg, malesame_mt_avg, maledown_mt_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

@@

# Create a bar graph using ggplot
ggplot(males_mt, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent", fill = "Category") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

# FIGURE 1D: During Semester Results FEMALE Bar Graph

# Selecting data of grades during the semester for female students 
female_mt_dta <- inst_data |>
              select(femaleup_mt, femalesame_mt, femaledown_mt)

# Calculating average of grade increase, no change, and decrease for female students 
femaleup_mt_avg <- mean(female_mt_dta$femaleup_mt, na.rm = TRUE)
femalesame_mt_avg <- mean(female_mt_dta$femalesame_mt, na.rm = TRUE)
femaledown_mt_avg <- mean(female_mt_dta$femaledown_mt, na.rm = TRUE)

# Creating new data frame with selected data
females_mt <- data.frame(
  value = c(femaleup_mt_avg, femalesame_mt_avg, femaledown_mt_avg),
  category = c("Grade Increase", "No Change", "Grade Decrease")
)

@@

# Create a bar graph using ggplot
ggplot(females_mt, aes(x = category, y = value, fill = category)) + ylim(0, 90) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(value)), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent", fill = "Category") + scale_fill_manual(values = c("Grade Increase" = "#4B659D", "No Change" = "#A83B48", "Grade Decrease" = "#B7BA9F")) +  # Adjust colors
  scale_y_continuous(breaks = custom_breaks) +  # Set custom breaks for y-axis
  theme_minimal() + theme(legend.position = "bottom")

@@

@fig-RegradeByClass shows the averaged distribution of the number of classes in which students are considered for regrades by gender. Observable in the graph, there is only data available on students who considered regrades in ‘0 class’ to ‘5 classes or more’ ([@fig-RegradeByClass]). Therefore, with the number of classes limited to 5, it’s more than likely that surveys understate the regraded data [@GenderDiff]. Observably, with no or 0 classes, the majority of students didn’t consider regrade requests, with females upholding 41.2 % while males upholding 39.3%, making it a 1.9% difference. This trend of females considering for regrades is higher than males when the class number is lower, where at 1 class, 30.4% of females are considered for regrades in contrast to  22.7% of males. This translates to a huge margin of 7.7% between the two genders.

However, as the number of classes extended above 1, or from 2 to 5 or more classes, more males considered asking for a regrade. For instance, at 3 classes, 8.9% of males considered for regrades as opposed to 7.1% of females; at 4 classes, 3.3% of males considered for regrades compared to 2% of females and finally, at 5 classes or more, a great 5.2% of males considered for regrades compared to a small 2.1% females. Statistically then, while the difference is negligible in some categories of the class number, males are more likely to consider and ask for regrades compared to females (except for those with 1 class).

@@

# Discussion {#sec-discussion}

The original paper looked to analyze the gender differences among undergraduate students upon regrade requests. The main question Li and Zahar analyzed was whether these requests result in favourable outcomes for either gender. To extrapolate on the original findings, we look to understand if the gender differences in regrade requests also correlate with the gender identity of the regrading instructors.

@@

One key point to consider is the disparity in outcomes between male and female students when requesting a regrade. The data shows that a higher percentage of male students (62%) received no change in their grade after requesting a regrade during the semester compared to female students (59%). This explains the notion that males typically send out higher requests of regrades compared to their female counterparts. This could be attributed to the social implications placed on each gender, confirming the concept that men are more assertive hence their inclination to ask for regrades. Women are more accepting hence they refrain from attempting to submit a request [@AACSB]. When looking at the percentage of students who received an increase in their grade after requesting a regrade, it turns out that a higher proportion of female students (40%) experienced this outcome than male students (37%). This suggests that, although male students were less successful overall in achieving grade changes when changes did occur, they were slightly more likely to increase for female students. Statistically, this 3% difference in outcomes, does not explain a great deal regarding the skewed gender implications but further research is required to understand the broader implications.

```{r}
#| label: fig-RegreadeResultsByInstGenderFinal
#| fig-cap: Instructor Survey End of Semester Regrade Results by Gender of Students and Gender of Instructor
#| fig-subcap: ["Panel A.  End of Semester Grades, Male Instructors", "Panel B. End of Semester Grades, Female Instructors"]
#| layout-ncol: 1
#| warning: false
#| message: false
#| echo: false

# Based off code from Chat GPT 3.5 
# PANEL A: During Semester Grades, Male Instructors Bar Graph

# Gathering Data of End of Semester Grades for Male and Female Students who had a Male Instructor
male_inst_dta <- inst_data_cleaned |>
  filter(inst_gender == "Male")

# Male Students
maleup_avg_maleInst <- mean(male_inst_dta$maleup_fn, na.rm = TRUE)
malesame_avg_maleInst <- mean(male_inst_dta$malesame_fn, na.rm = TRUE)
maledown_avg_maleInst <- mean(male_inst_dta$maledown_fn, na.rm = TRUE)

# Female Students
femaleup_avg_maleInst <- mean(male_inst_dta$femaleup_fn, na.rm = TRUE)
femalesame_avg_maleInst <- mean(male_inst_dta$femalesame_fn, na.rm = TRUE)
femaledown_avg_maleInst <- mean(male_inst_dta$femaledown_fn, na.rm = TRUE)

# Create a data frame with the results
inst_results_male <- data.frame(
  value = c(maleup_avg_maleInst, femaleup_avg_maleInst, malesame_avg_maleInst, femalesame_avg_maleInst, maledown_avg_maleInst, femaledown_avg_maleInst),
  category = rep(c("Grade Increase", "No Change", "Grade Decrease"), each = 2),
  student = rep(c("Male", "Female"), times = 3)
)

# Convert the 'Category' variable to a factor with the specified order
inst_results_male$category <- factor(inst_results_male$category, levels = category_order)

# Create a dodged bar graph using ggplot with position_dodge2
ggplot(inst_results_male, aes(x = category, y = value, fill = student)) +
  geom_bar(stat = "identity", position = position_dodge2(width = 0.9), width = 0.7) +
  geom_text(aes(label = round(value)), position = position_dodge2(width = 0.9), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent", 
       fill = "Student Gender") +
  scale_fill_manual(values = c("Male" = "#4B659D", "Female" = "#A83B48")) +  # Adjust colors
  theme_minimal()

# PANEL B: During Semester Grades, Female Instructors Bar Graph

# Gathering Data of End of Semester Grades for Male and Female Students who had a Female Instructor
male_inst_dta <- inst_data_cleaned |>
  filter(inst_gender == "Female")

# Male Students
maleup_avg_maleInst <- mean(male_inst_dta$maleup_fn, na.rm = TRUE)
malesame_avg_maleInst <- mean(male_inst_dta$malesame_fn, na.rm = TRUE)
maledown_avg_maleInst <- mean(male_inst_dta$maledown_fn, na.rm = TRUE)

# Female Students
femaleup_avg_maleInst <- mean(male_inst_dta$femaleup_fn, na.rm = TRUE)
femalesame_avg_maleInst <- mean(male_inst_dta$femalesame_fn, na.rm = TRUE)
femaledown_avg_maleInst <- mean(male_inst_dta$femaledown_fn, na.rm = TRUE)

# Create a data frame with the results
inst_results_male <- data.frame(
  value = c(maleup_avg_maleInst, femaleup_avg_maleInst, malesame_avg_maleInst, femalesame_avg_maleInst, maledown_avg_maleInst, femaledown_avg_maleInst),
  category = rep(c("Grade Increase", "No Change", "Grade Decrease"), each = 2),
  student = rep(c("Male", "Female"), times = 3)
)

# Convert the 'Category' variable to a factor with the specified order
inst_results_male$category <- factor(inst_results_male$category, levels = category_order)

# Create a dodged bar graph using ggplot with position_dodge2
ggplot(inst_results_male, aes(x = category, y = value, fill = student)) +
  geom_bar(stat = "identity", position = position_dodge2(width = 0.9), width = 0.7) +
  geom_text(aes(label = round(value)), position = position_dodge2(width = 0.9), vjust = -0.5, size = 3) +  # Add labels on top of bars
  labs(x = "Category",
       y = "Percent", 
       fill = "Student Gender") +
  scale_fill_manual(values = c("Male" = "#4B659D", "Female" = "#A83B48")) +  # Adjust colors
  theme_minimal()
